\section{Couplers \label{couplers}}

The action in \module{httpy} is in the responders, but the couplers are the
workhorses that put responders on the network. \module{httpy} includes three
couplers out of the box. There are multiple Python webservers available, and a
couple Python FastCGI libraries too. I've tried to pick the single most mature
option in each category, and build the couplers on top of those. These decisions
are purely pragmatic.

\begin{classdesc}{CGI}{Responder} Constructs a new \class{CGI} coupler object.
This class is defined in the \module{httpy.couplers.cgi} module.
\end{classdesc}

\begin{classdesc}{FastCGI}{Responder} Constructs a new \class{FastCGI} coupler
object. This class is defined in the \module{httpy.couplers.fastcgi} module.
\end{classdesc}

\begin{classdesc}{StandAlone}{Responder\optional{, argv}} Constructs a new
\class{StandAlone} coupler object. \var{argv} is a sequence of arguments. If
omitted or \class{None}, \code{sys.argv} is used. See \ulink{the chapter on the
\program{httpy} executable}{manual.html} for the available options. This class
is defined in the \module{httpy.couplers.standalone} module. \end{classdesc}

Each coupler takes a class defining a (possibly implicit) provider of the
\class{IResponder} interface as its first argument, and each provides the
following method:

\begin{methoddesc}{go}{} For \class{CGI} instances, this responds to a single
request. For \class{FastCGI} and \class{StandAlone}, this enters a blocking
loop.\end{methoddesc}

As an example, here is what a CGI script looks like with \module{httpy}:

\begin{verbatim}
#!/usr/local/bin/python
"""This is a CGI script.
"""
from httpy.couplers.cgi import CGI
import MyResponder

coupled = CGI(MyResponder)
coupled.go()
\end{verbatim}

Those wishing to implement new couplers are invited to consult the source code,
and to contact the author.