\section{Responders \label{responders}}

There are basically two kinds of websites, publications and applications,
differentiated by their organization and interface models. A \dfn{publication}
website organizes information into individual pages within a hierarchical folder
structure that one navigates by browsing. In an \dfn{application} website, on
the other hand, data is not organized into hierarchical pages but is dealt with
via a non-browsing interface such as a search box.

The HTML version of this documentation is an example of a publication website: a
number of hypertext documents organized into sections. If we weren't using LaTeX
(or if I knew how to use it better), the sections would probably be encoded in
folders. \ulink{Gmail}{http://mail.google.com/mail} is a pure application
website, one which organizes and presents information non-hierarchically. Most
websites, however, are hybrids. That is, within an overall hierarchical
organization you will find both individual pages of information as well as
applications such as a site search feature, or a threaded discussion forum.

Publication websites are actually a subset of application websites, of course.
An application site can use any interface metaphor; a publication is an
application that uses the familiar folder/page metaphor to organize and present
its information. Therefore, every website is fundamentally an application.

\module{httpy} enables you to build custom hybrid, publication, and application
websites, and its bundled responders include an implementation of each.




\begin{classdesc}{Multiple}{\optional{root}} Constructs a new \class{Multiple}
responder object. \var{root} is the filesystem path of the directory to use as
the publishing root. If not given or \class{None}, then the value in the class
attribute of the same name is used if it does not evaluate to \class{False}. If
it does (and out of the box, it does), then \var{root} is set to the current
working directory. If \var{root} does not point to a directory, \exception
{ValueError} is raised.\end{classdesc}

\begin{classdesc}{Static}{\optional{root}\optional{, defaults}} Constructs a new
\class{Static} responder object. \var{root} has the same meaning as in
\class{Multiple}. \var{defaults} is a sequence of names to use when looking for
a default resource. If it is omitted or \class{None}, it is set to
\code{('index.html', 'index.htm')}, after checking the class attribute as with
\var{root}. \end{classdesc}

\begin{classdesc*}{XMLRPC}
The XMLRPC responder is not designed to be instantiated directly, but rather
used as a mixin.
\end{classdesc*}


\input{multiple}
\input{static}
\input{xmlrpc}
\input{responder}
