\subsection{\class{XMLRPC} Responders \label{xmlrpc}}

\ulink{XMLRPC}{http://www.xmlrpc.com/} is a language-neutral protocol for
distributed computing, using XML for serialization and HTTP for transport. The
Python standard library includes \ulink{a well-written XMLRPC
library}{http://docs.python.org/lib/module-xmlrpclib.html}, which forms the
basis for this class. By mixing the \class{XMLRPC} responder into your class,
you can easily make your instance's methods available to clients over the
network. Instances inherit one attribute and two methods:

\begin{memberdesc}[tuple]{protected} A tuple of names of methods of \class{self}
which should never be served via XMLRPC. This class attribute on \class{XMLRPC}
is set to the empty tuple.\end{memberdesc}

\begin{methoddesc}{respond}{request}
This is a pass-through for \method{serve_xmlrpc}.
\end{methoddesc}

\begin{methoddesc}{serve_xmlrpc}{request}
Proxies methods of \class{self} via XMLRPC. \var{request} is a \class{Request}
object. Assuming \var{request} is an XMLRPC request for method \var{name}, the
following conditions trigger an XMLRPC <fault> response with <faultCode> 404:

\begin{enumerate}

\item \var{name} starts with an underscore.

\item \var{name} is \code{respond} or \code{serve_xmlrpc}.

\item \var{name} is named in \code{protected}.

\item The instance does not have a method named \var{name}.

\end{enumerate}
\end{methoddesc}


\subsubsection{An Example}

In a Python shell, create an XMLRPC server like so:

\begin{verbatim}
>>> import httpy
>>> class Responder(httpy.responders.XMLRPC):
...   protected = ['private']
...   def private(self):
...     return 'leave me alone!'
...   def ping(self):
...     return 'pong'
...
>>> responder = Responder()
>>> coupler = httpy.couplers.StandAlone(responder)
>>> coupler.go()
httpy.server     INFO     httpy started on port 8080
\end{verbatim}

Then, in a second shell, you can talk to your server like this:

\begin{verbatim}
>>> import xmlrpclib
>>> server = xmlrpclib.ServerProxy('http://localhost:8080/')
>>> server.ping()
'pong'
>>> server.private()
Traceback (most recent call last):
...
xmlrpclib.Fault: <Fault 404: "method 'private' not found">
>>>
\end{verbatim}
