\section{\class{Request} Objects \label{request}}

HTTP is a stateless protocol. Any application state that exists in a request is
encoded in one of the following places:

\begin{itemize}
\item{the Request Method}
\item{the path}
\item{a querystring}
\item{another part of the Request-URI}
\item{a \mailheader{Cookie} header}
\item{another header}
\item{a \code{POST} body}
\end{itemize}

httpy doesn't pretend to know the best way for you to encode state in your app's
requests, nor the best way to object-represent this state. Instead, it gives you
the raw HTTP message, and a very minimal secondary API. There are a few tools in
\class{httpy.utils} that might be helpful -- \function{parse_body},
\function{parse_cookie}, \function{parse_query} -- but you are free to translate
the \class{Request} to your application's object model however you like.


\subsection{Raw API}

Instances of \class{httpy.Request} store the raw HTTP Request message in the
following attributes:

\begin{datadesc}{raw}
The entire \ulink{HTTP Request message}{http://www.w3.org/Protocols/rfc2616/rfc2616-sec5.html} exactly as it was found on the wire.
\end{datadesc}

\begin{datadesc}{raw_line}
The raw
\ulink{Request-Line}{http://www.w3.org/Protocols/rfc2616/rfc2616-sec5.html#sec5.1},
 not including the trailing line break.
\end{datadesc}

\begin{datadesc}{raw_headers}
The raw message headers, not including the trailing line breaks.
\end{datadesc}

\begin{datadesc}{raw_body}
The raw message body.
\end{datadesc}



\subsection{Derived API}

Besides providing access to the raw Request message, \class{Request} instances
provide a very minimal derivative API:

\begin{datadesc}{method}
The \ulink{HTTP method}{http://www.w3.org/Protocols/rfc2616/rfc2616-sec9.html} from the Request-Line.
\end{datadesc}

\begin{datadesc}{uri}
The \ulink{Request-URI}{http://www.w3.org/Protocols/rfc2616/rfc2616-sec5.html#sec5.1.2} as a dictionary. The keys of this dictionary are taken from the
names used in the standard library's
\ulink{\module{urlparse}}{http://docs.python.org/lib/module-urlparse.html} module,
namely: \emph{scheme://netloc/path;parameters?query\#fragment}.
\end{datadesc}

\begin{datadesc}{path}
The \emph{path} component of the Request-URI.
\end{datadesc}

\begin{datadesc}{headers}
The headers as an instance of the standard library's
\ulink{\module{email.Message.Message}}{http://docs.python.org/lib/module-email.Message.html}.
\end{datadesc}


\subsection{Differences under CGI/FastCGI}

Whereas the \class{StandAlone} coupler builds \class{Request} objects directly
from the raw request coming off the wire, the \class{CGI} and \class{FastCGI}
couplers must resort to reconstructing \class{Request}s from the environment and
standard input. The resulting object has the same API as above, but since some
attributes could have subtly different meanings, we provide the following
reference. Where called for, \code{\e r\e n} line breaks are used.

\begin{datadesc}{raw}
The concatenation of \member{raw_line}, \member{raw_headers}, and
\member{raw_body} as described below.
\end{datadesc}

\begin{datadesc}{raw_line}
The Request-URI is reconstructed from the \envvar{SCRIPT_NAME},
\envvar{PATH_INFO}, and \envvar{QUERY_STRING} environment variables. This is
combined with the \envvar{REQUEST_METHOD} and \envvar{SERVER_PROTOCOL} variables
to approximate the Request-Line.
\end{datadesc}

\begin{datadesc}{raw_headers}
The headers are reconstructed by taking all environment variables beginning with
"HTTP_" and replacing all underscores with dashes.
\end{datadesc}

\begin{datadesc}{raw_body}
Read from the standard input.
\end{datadesc}

\begin{datadesc}{method}
Equivalent to the \envvar{REQUEST_METHOD} environment variable.
\end{datadesc}

\begin{datadesc}{uri}
The dictionary is based on \member{raw_line} as derived above.
\end{datadesc}

\begin{datadesc}{path}
Equivalent to the \envvar{PATH_INFO} environment variable.
\end{datadesc}

\begin{datadesc}{headers}
The \class{Message} object is constructed from \member{raw_headers} as described
above.
\end{datadesc}


\begin{seealso}
    \seetitle[http://hoohoo.ncsa.uiuc.edu/cgi/interface.html]{The CGI
    Specification}{ The reference for CGI, including use of environment
    variables and standard input.}
\end{seealso}
